Large-scale spatiotemporal variation in vital rates and population dynamics of an alpine bird

### 2024-07-23 Manuscript_Nater_etal.qmd
Addition to introduction

There is growing demand for biodiversity indicators from international unions, national governments, local management bodies, and corporate and industry actors. Indicators should ideally represent a wide range of biodiversity's states and functions [e.g. Essential Biodiversity Variables, @pereira2013; @jetz2019], yet the development of suitable indicators for certain attributes, such as species abundance and demography, has been more difficult than for others [@schmeller2018; @waldock2022]. This is at least partially due to challenging requirements regarding spatial scales of useful biodiversity indicators. On one hand, indicators need to be representative at large geographic scales, for example, in the context of countries' reporting towards biodiversity targets [e.g. @feld2009]. On the other hand, indicators also ideally have good spatial resolution, as the scales relevant for local-level management and planning are often much smaller [@stevenson2021]. This latter requirement is particularly crucial for infrastructure development strategies and for species management and conservation, both of which tend to require knowledge on species abundance and population dynamics (i.e. demographic rates) that is relevant for county- or municipality-level decision making [@christie2020]. Another reason why abundance and population indicators ideally come with good spatial resolution is that there can be substantial amounts of variation in population dynamics and life history of species across space [e.g. @robinson2014; @horswill2019]. This variation needs to be accounted for to develop successful and sustainable strategies for area use, harvest management, and species and biodiversity conservation [@williams2002].

>>>>

There is growing demand for biodiversity indicators from international unions, national governments, local management bodies, and corporate and industry actors. Indicators should ideally represent a wide range of biodiversity's states and functions [e.g. Essential Biodiversity Variables, @pereira2013; @jetz2019], yet the development of suitable indicators for certain attributes, such as species abundance and demography, has been more difficult than for others [@schmeller2018; @waldock2022]. This is at least partially due to challenging requirements regarding spatial scales of useful biodiversity indicators. On one hand, indicators need to be representative at large geographic scales, for example, in the context of countries' reporting towards biodiversity targets [e.g. @feld2009]. On the other hand, indicators also ideally have good spatial resolution, as the scales relevant for local-level management and planning are often much smaller [@stevenson2021]. This latter requirement is particularly crucial for infrastructure development strategies and for species management and conservation, both of which tend to require knowledge on species abundance and population dynamics (i.e. demographic rates) that is relevant for county- or municipality-level decision making [@christie2020]. Another reason why abundance and population indicators ideally come with good spatial resolution is that there can be substantial amounts of variation in population dynamics and life history of species across space [e.g. @robinson2014; @horswill2019]. Such variation can arise from differences in ecological contexts (including local habitat and weather conditions, hunting pressures, and interspecific interactions, e.g. @nilsen2009; @bond2021) and needs to be accounted for when developing sustainable strategies for area use, harvest management, and species and biodiversity conservation [@williams2002].
